The population of long-period exoplanets --- \\ Fully-automated discovery \& characterization of long-period transiting exoplanet candidates in the \kepler\ archive

### 2016-04-19 document/ms.tex
words

% \documentclass[twocolumn]{aastex6}
\documentclass[onecolumn]{aastex6}
% \documentclass[preprint]{aastex}

@@

\title{%
The population of long-period exoplanets I. --- \\
Fully-automated discovery \& characterization of
long-period transiting exoplanet candidates in the \kepler\ archive
}

@@

There are two main technical barriers to a search for single transit events.
The first is that the transit probability for long-period planets is very low;
scaling as $\propto P^{-5/3}$ for orbital periods longer than the
baseline of contiguous observations.
Therefore, even if long-period planets are intrinsically common, they will
still be underrepresented in a transiting sample.
The second challenge is that there are substantial signals in the observed
light curves caused by stochastic processes---both instrumental
(pointing jitter, temperature variations, \etc) and astrophysical (stellar
variability, \etc)---that can masquerade as transit signals.
In practice, even using the most sophisticated systematics removal methods,
these false signal far outnumber the true single transits.

@@

A qualitatively different approach to planet search is employed by the
\project{Planet Hunters} (PH)
project\footnote{\url{http://www.planethunters.org/}} \citep{Fischer:2012}.
PH is a ``citizen science'' project where visitors to the website look at
sections of light curve and mark the locations of transits that can be
identified visually.
Visual inspection can be a useful search technique for large single transits
because humans are able to robustly distinguish transit signals from the noise
and this project has yielded some promising long-period candidates and
confirmed planets \citep[for example][]{Wang:2013} \dfmtodo{CITE Wang 2015}.
One shortcoming of the PH method is that it can be difficult to fully quantify
the performance (completeness and reliability) of the search and PH does not
evaluate their users using synthetic transit signals as is now common practice
in the transit search literature \dfmtodo{is this still true?}.
This means that the PH sample of candidates cannot be used for robust
population inference.

To study the occurrence rate of long-period exoplanets, we have developed a
fully-automated method for searching the archival \kepler\ light curves for
these systems with only one or two transits.
Using this method, we detect XXX long-period transiting planet candidates and
YYY long-period eclipsing binary candidates.
We estimate that the population level false positive rate in this exoplanet
candidate sample is about ZZZ\%.
By injecting synthetic transit signals into the \kepler\ light curves, we
measure the completeness of our search method.
Based on this catalog of exoplanet candidates and the search completeness, we
constrain the occurrence rate of long-period exoplanets.
For large planets, these results are consistent with previous measurements of
long-period, giant planet occurrence rates based on radial velocity surveys
and we demonstrate that small---and possibly rocky---planets are more common
than their larger counterparts even at long orbital periods.

Data from the \kepler\ Mission has been used to discover a diverse zoo of
thousands of exoplanets of all sizes orbiting stars of all types.
It has be demonstrated that these catalogs of exoplanet discoveries can be
used to place constraints on the occurrence rate of exoplanets \dfmtodo{CITE}.
These results are sensitive to the detection efficiency and reliability of the
discovery methodology \dfmtodo{CITE Burke}.

@@

@@

It is not feasible to compute a full model comparison at every time in the
light curve so we must first find potentially interesting events.
To generate this list, we apply a standard ``box least squares'' (BLS;
\dfmtodo{CITE}) method with a single (non-periodic) box.
First, we filter the PDC (\dfmtodo{CITE}) light curves using a running
windowed median with a half-width of \dfmtodo{how many} to remove stellar
variability.
We then compute the signal-to-noise of the depth of a 0.6~day long transit on
a grid of times spanning the full \kepler\ baseline.
To avoid edge effects, we apodize this detection scalar near any large gaps in
the time series using a logistic function with width equal to one transit
duration.
Finally, we estimate the noise in the time series using a robust running
windowed variance estimate of the detection scalar and accepting peaks more
than 25-times this noise as candidates.

@@

>>>>

% \documentclass[twocolumn]{aastex6}
% \documentclass[onecolumn]{aastex6}
\documentclass[preprint,12pt]{aastex}

@@

\title{%
The population of long-period exoplanets --- \\
Fully-automated discovery \& characterization of
long-period transiting exoplanet candidates in the \kepler\ archive
}

@@

There are two main technical barriers to a search for single transit events.
The first is that the transit probability for long-period planets is very low;
scaling as $\propto P^{-5/3}$ for orbital periods longer than the
baseline of contiguous observations.
Therefore, even if long-period planets are intrinsically common, they will
still be underrepresented in a transiting sample.
The second challenge is that there are many signals in the observed
light curves caused by stochastic processes---both instrumental
(pointing jitter, temperature variations, \etc) and astrophysical (stellar
variability, \etc)---that can masquerade as transits.
Even using the most sophisticated methods for removing this variability, false
signal far outnumber the true single transits.

@@

A qualitatively different approach to planet search is employed by the
\project{Planet Hunters} (PH)
project\footnote{\url{http://www.planethunters.org/}} \citep{Fischer:2012}.
PH is a ``citizen science'' project where visitors to the website look at
sections of light curve and mark the locations of transits that can be
identified visually.
Visual inspection can be a useful search technique for large single transits
because humans are able to robustly distinguish transit signals from the noise
and this project has yielded some promising long-period candidates and
confirmed planets \citep[for example][]{Wang:2013} \dfmtodo{CITE Wang 2015}.
One shortcoming of the PH method is that it can be difficult to fully quantify
the performance (completeness and reliability) of the search results.
% and PH does not
% evaluate their users using synthetic transit signals as is now common practice
% in the transit search literature \dfmtodo{is this still true?}.
This means that the PH sample of candidates cannot be reliably used for robust
population inference.

Visual inspection and hand-curation by experts has also been used to find the
transits of long-period planets in previously known planetary systems
(\dfmtodo{CITE single transit KOIs}).
All light curves in the \kepler\ archive with \kepler\ Objects of Interest
(KOIs) were visually inspected to find candidate single transits.
This is a reasonable idea because long-period planets with transiting inner
planets are expected to have a higher geometric transit probability.
The selection effects introduced by restricting the target sample like this
are difficult to take into account when making inferences about the population
of long-period planets.

To study the occurrence rate of long-period exoplanets, we have developed a
fully-automated method to search the archival \kepler\ light curves and
reliably discover long-period planets with only one or two transits.
Using this method, we search the 40,000 brightest Sun-like stars and detect
XXX long-period astrophysical transit candidates.
Of these, YYY are likely to be exoplanets with radii equal to or smaller than
Jupiter's radius.
Using transit injection and recovery, we measure the completeness of our
search procedure.
We estimate that the population level astrophysical false positive rate in
this candidate sample is about ZZZ\%.
By injecting synthetic transit signals into the \kepler\ light curves, we
measure the completeness of our search method.
Based on this catalog of exoplanet candidates and the search completeness, we
constrain the occurrence rate of long-period exoplanets.
For large planets, these results are consistent with previous measurements of
long-period, giant planet occurrence rates based on radial velocity surveys
and we demonstrate that small---and possibly rocky---planets are more common
than their larger counterparts even at long orbital periods.

@@

The model comparison step is computationally expensive because we estimate the
marginalized likelihoods for the light curve data under several different
models describing a transit and many other processes that ``look'' like
transits but are actually caused by noise.
This step is also conservative; unless a signal is a very convincing transit,
it won't pass the test.
In practice, this means that all but the highest signal-to-noise events will
be rejected at this step.
Therefore, in the inexpensive first step---the initial candidate search---we
can restrict the candidate list to high signal-to-noise events without a
substantial loss in detection efficiency.

@@

It is not computationally feasible to run a full model comparison at every
time in the light curve so we must first find potentially interesting events.
For our purposes, ``interesting'' means high signal-to-noise and previously
unknown.
We restrict our search to high signal-to-noise transits since the

To generate this list, we apply a standard ``box least squares'' (BLS;
\dfmtodo{CITE}) method with a single (non-periodic) box.
First, we filter the PDC (\dfmtodo{CITE}) light curves using a running
windowed median with a half-width of \dfmtodo{how many} to remove stellar
variability.
We then compute the signal-to-noise of the depth of a 0.6~day long transit on
a grid of times spanning the full \kepler\ baseline.
We only need to use a single transit duration because the following steps in
this procedure are only sensitive to transits with very high signal-to-noise
so, in practice, the results of this step are insensitive to the specific
choice of duration.

As previously mentioned, the following steps will only detect high
signal-to-noise transits.
Therefore, we use this
To avoid edge effects, we apodize this detection scalar near any large gaps in
the time series using a logistic function with width equal to one transit
duration.
Finally, we estimate the noise in the time series using a robust running
windowed variance estimate of the detection scalar and accepting peaks more
than 25-times this noise as candidates.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/completeness.pdf}
\end{center}
\caption{%
\figlabel{completeness}}
\end{figure*}
